### 2019-01-22 R/README.Rmd
Update each README.md
Update each README.md to describe contents of each directory.

Remove CONDUcT.md and CONTRIBUTION.md.

Add the following:
 - mcp.csl file for correct citation style.

 - pagebreak.lua, a filter to convert LaTeX \newpage command into
   appropriate markup for other formats e.g. html, docx. See:
   https://github.com/pandoc/lua-filters/tree/master/pagebreak

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R

This directory contains R script files used throughout the project.

Each file contains several custom functions which have been
documented using roxygen2 skeletons.

# contents
```
    .
    ├── ...
    ├── /R
    │    ├── file1.R          # Functions for x
    │    ├── file2.R          # Functions for y
    │    ├── file3.R          # Functions for z
    │    ├── README.Rmd       # R markdown file to produce README.md
    │    └── README.md        # Description of /R directory
    └── ...

```