### 2019-01-22 analysis/README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# analysis

This directory contains the complete data analysis workflow upon which the paper

> Authors, (YYYY). _Title of paper_. Name of journal/book <https://doi.org/xxx/xxx>

is based upon.

# contents
```
    .
    ├── ...
    ├── /analysis
    │    ├── /figures           # Contains figures generated by analysis R scripts
    │    ├── analysis.Rmd       # R markdown file to produce analysis.html/analysis.md (haven't decided)
    │    ├── analysis.html      # Contains cleaned reference tables
    │    ├── analysis.md        # Contains final results tables
    │    ├── file1.R            # Script for x
    │    ├── README.Rmd         # R markdown file to produce README.md
    │    └── README.md          # Description of /analysis directory
    └── ...

```