### 2019-01-22 manuscript/README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# manuscript

This directory contains the final manuscript as submitted to Molecular
and Cellular Proteomics. It also contains templates and other files necessary 
for producing the `manuscript.pdf` file from the associated `manuscript.Rmd`.

# contents
```
    .
    ├── ...
    ├── /manuscript
    │    ├── /figures            # Contains figures draw by hand
    │    ├── /templates          # Contains citation and pagebreak templates
    │    ├── manuscript.Rmd      # R markdown file to produce manuscript.pdf
    │    ├── manuscript.pdf      # Final submitted manuscript
    │    ├── references.bib      # BibTex file containing manuscript references
    │    ├── README.Rmd          # R markdown file to produce README.md
    │    └── README.md           # Description of /manuscripts directory
    └── ...

```