### 2019-02-13 data-raw/README.Rmd
Move raw data
 - Begin restructure of repository to conform to traditional structure
   of an R package.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# data

This directory contains all of the data files used to generate results
tables and figures.

# contents
```
    .
    ├── ...
    ├── /data
    │    ├── /raw            # Contains raw untouched data files
    │    ├── /external       # Contains data from other papers
    │    ├── /clean          # Contains cleaned reference tables
    │    ├── /results        # Contains final results tables
    │    ├── README.Rmd      # R markdown file to produce README.md
    │    └── README.md       # Description of /data directory
    └── ...

```