### 2019-02-13 data-raw/README.Rmd
Update ./data-raw README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# /data-raw

## description
This directory contains all of the raw data files used in this project. These files are
*not* included when the package is installed using `devtools::install_github()`. Rather 
*'clean'* versions of the `.csv` and `.tab` files are provided in the [`/data`](../data) 
folder as `.rda` files which are accessible once the package has been installed and 
loaded or attached.

To access these raw files:

1. Right click on a desired file and select `Save link as...`.

2. Copy the url for the desired directory, paste into 
[DownGit](https://minhaskamal.github.io/DownGit/#/home), then download as a `.zip`.

3. Download this entire repository as a `.zip` from this URL: 
<https://github.com/csdaw/candidaev/archive/master.zip>

4. Clone this repository.

## contents
```
    .
    ├── ...
    ├── /data-raw
    │    ├── /external       # Supplementary data tables from other Candida EV proteomics papers
    │    ├── /lcms           # proteinGroups.txt files output by MaxQuant
    │    ├── /nta            # Nanosight ns300 .csv summary tables
    │    ├── /qubit          # Qubit 4 EV protein quantification data 
    │    ├── /reference      # UniProt and Candida Genome Database reference tables
    │    ├── README.Rmd      # R markdown file to produce README.md
    │    └── README.md       # Description of /data-raw directory
    └── ...

```