### 2019-06-05 inst/manuscript/README.Rmd
Move manuscript directory to inst/

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# /vignettes

## description
This directory contains the submitted manuscript (see `manuscript.pdf`) 
for this paper :

> Authors, (YYYY). _Title of paper_. Name of journal/book <https://doi.org/xxx/xxx>

It also contains a more detailed analysis of the LC-MS/MS data, upon which the manuscript 
is based (see `analysis.md` or `analysis.html`). The 
Rmarkdown (`.Rmd`) files used to generate these documents are provided.

Both the manuscript and the analysis documents *are* included as vignettes in this 
package so they can be easily viewed (or regenerated completely from the raw data) if 
the package is installed. Alternatively, one can download these documents directly 
from this GitHub repository.

## how to access the manuscript

## how to access the analysis

## contents
```
    .
    ├── ...
    ├── /vignettes
    │    ├── /figures            # Contains figures
    │    ├── /components         # Contains required components for knitting manuscript and analysis
    │    ├── analysis.Rmd        # R markdown file to produce analysis.md or analysis.html
    │    ├── analysis.html       # Full analysis upon which the manuscript is based.
    │    ├── analysis.md         # Full analysis upon which the manuscript is based.
    │    ├── manuscript.Rmd      # R markdown file to produce manuscript.pdf or manuscript.docx
    │    ├── manuscript.pdf      # Final submitted manuscript
    │    ├── README.Rmd          # R markdown file to produce README.md
    │    └── README.md           # Description of /vignettes directory
    └── ...

```