### 2019-06-18 inst/manuscript/manuscript_supplement.Rmd
Add ribosome section to supplemental Table S2

---
title: | 
   **Defining protein markers for *Candida albicans* extracellular vesicles**
author: | 
   *Charlotte Dawson^a^, Mark Bleackley^a^, Donovan Garcia Ceron^a^, 
   Marilyn Anderson^a^*
date: | 
   *^a^Department of Biochemistry and Genetics, La Trobe Institute for 
   Molecular Science, La Trobe University, Victoria, 3086, Australia*
output: 
    bookdown::pdf_document2:
      latex_engine: xelatex
      fig_caption: yes
      number_sections: yes
      toc: no
      keep_tex: no
bibliography: "components/references.bib" 
csl: "components/molecular-and-cellular-proteomics.csl"
link-citations: yes
linkcolor: black
fontsize: 12pt
geometry: margin=1in
classoption: letterpaper
mainfont: Times New Roman
linestretch: 1.5
header-includes:
  - \usepackage{titling}
  - \pretitle{\begin{center}\fontsize{14bp}{14bp}\selectfont}
  - \posttitle{\par\end{center}}
  - \preauthor{\begin{center}\fontsize{12bp}{12bp}\selectfont}
  - \postauthor{\par\end{center}}
  - \predate{\begin{center}\fontsize{10bp}{10bp}\selectfont}
  - \postdate{\par\end{center}}
  - \usepackage{fancyhdr}
  - \usepackage{lipsum}
  - \pagestyle{fancy}
  - \fancyhead[L]{This is the running title it is rather long}
  - \setlength{\headheight}{15pt}
  - \usepackage{float}
  - \let\origfigure\figure
  - \let\endorigfigure\endfigure
  - \renewenvironment{figure}[1][2]{\expandafter\origfigure\expandafter[H]}
    {\endorigfigure}
  - \newcommand{\beginsupplement}{
      \setcounter{table}{0} 
      \renewcommand{\thetable}{S\arabic{table}} 
      \setcounter{figure}{0} 
      \renewcommand{\thefigure}{S\arabic{figure}}
    }
  - \usepackage{caption}
  - \captionsetup{font={normalsize,stretch=1.3},labelfont=bf}
  - \usepackage{threeparttable}
---

**Running title:** Running title goes here

**Corresponding author:**

\newpage

```{r setup, include=FALSE}
# set knit options
knitr::opts_knit$set(progress = TRUE, 
                     verbose = TRUE)

# set chunk options
knitr::opts_chunk$set(collapse = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      echo = FALSE, 
                      comment = "#>")

# load libraries
library(candidaev)
library(tidyNano)
library(limma)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(kableExtra)

# set seed for random number generation
set.seed(1)

# reference external R script with protein function lists
knitr::read_chunk('components/protein_funs.R')
```

```{r yeast, include=FALSE}
# filter out reverse, contaminant proteins, proteins with <2 unique peptides
# extract LFQ intensity columns and UniProt accessions 
# log2 transform LFQ intensities
yeast_lfq <- yeast %>% 
  filter(Reverse != "+", 
         Potential.contaminant != "+", 
         Unique.peptides >= 2) %>% 
  convert_lfq(., yeast_exp)

# drop EV_1 and WCL_1
yeast_lfq2 <- yeast_lfq[, !colnames(yeast_lfq) %in% c("EV_1", "WCL_1")]

# filter for proteins quantified in min 2/3 reps of EV or WCL
yeast_filt2 <- filter_na2(yeast_lfq2, logic = "or", op = "<=", 
                          pat1 = "EV", val1 = 1, 
                          pat2 = "W", val2 = 1)

# normalise LFQ intensities
yeast_norm2 <- normalizeCyclicLoess(yeast_filt2)

# filter for proteins identified in min 1/3 reps of EV
# define 'EV proteins'
y_ev <- filter_na(yeast_norm2, op = "<=", 
                  pat = "EV", val = 2)

# filter for proteins identified in min 1/3 reps of WCL
# define 'WCL proteins'
y_wcl <- filter_na(yeast_norm2, op = "<=", 
                   pat = "W", val = 2)

# can't impute with so many missing values
# separate EV/WCL exclusive proteins (don't impute) from 
# EV/WCL common proteins (impute)
yeast_excl <- filter_na2(yeast_norm2, logic = "or", op = "==", 
                         pat1 = "EV", val1 = 3, 
                         pat2 = "W", val2 = 3)

yeast_both <- filter_na2(yeast_norm2, logic = "and", op = "<=", 
                         pat1 = "EV", val1 = 2, 
                         pat2 = "W", val2 = 2)

# proteins with missing values tend to have lower intensity
# therefore proteins are MNAR, close to detection limit
# use left censored imputation method
yeast_imp <- impute_QRILC(yeast_both)

# recombine imputed proteins and non-imputed proteins in matrix
yeast_de <- rbind(yeast_excl, yeast_imp)

# see limma user guide section 9.2 for more info about DE
# create design matrix
y_samp <- data.frame(T = (rep(c("EV", "WCL"), each = 3)))

y_design <- stats::model.matrix(~ 0 + T, data = y_samp)
colnames(y_design) <- c("EV", "WCL")

# define sample comparisons of interest
y_contrasts <- c("EV - WCL")

# make all pair-wise comparisons between EV and WCL
# and perform limma::eBayes()
y_efit <- limma_eBayes(yeast_de, design = y_design, contrasts = y_contrasts)

# extract DE results
yeast_res <- get_results(efit = y_efit, 
                         mat = yeast_de, 
                         p_val = 0.01, 
                         lfc = 0, 
                         type = "individual")[[1]]
```

```{r biofilm, include=FALSE}
# filter out reverse, contaminant proteins, proteins with <2 unique peptides
# extract LFQ intensity columns and UniProt accessions 
# log2 transform LFQ intensities
biofilm_lfq <- biofilm %>% 
  filter(Reverse != "+", 
         Potential.contaminant != "+", 
         Unique.peptides >= 2) %>% 
  convert_lfq(., biofilm_exp)

# filter for proteins quantified in min 4/5 reps of EV or WCL
biofilm_filt <- filter_na2(biofilm_lfq, logic = "or", op = "<=", 
                           pat1 = "EV", val1 = 1, 
                           pat2 = "W", val2 = 1)

# normalise LFQ intensities
biofilm_norm <- normalizeCyclicLoess(biofilm_filt)

# filter for proteins identified in min 1/5 reps of EV
# define 'EV proteins'
b_ev <- filter_na(biofilm_norm, op = "<=", 
                  pat = "EV", val = 4)

# filter for proteins identified in min 1/5 reps of WCL
# define 'WCL proteins'
b_wcl <- filter_na(biofilm_norm, op = "<=", 
                   pat = "W", val = 4)

# can't impute with so many missing values
# filter for proteins with 4-5 NA values in EV or WCL (don't impute)
biofilm_excl <- filter_na2(biofilm_norm, logic = "or", op = ">=", 
                         pat1 = "EV", val1 = 4, 
                         pat2 = "W", val2 = 4)

# filter for proteins with min 3 valid values in EV and WCL (impute)
biofilm_both <- filter_na2(biofilm_norm, logic = "and", op = "<=", 
                         pat1 = "EV", val1 = 3, 
                         pat2 = "W", val2 = 3)

# proteins with missing values tend to have lower intensity
# therefore proteins are MNAR, close to detection limit
# use left censored imputation method
biofilm_imp <- impute_QRILC(biofilm_both)

# recombine imputed proteins and non-imputed proteins in matrix
biofilm_de <- rbind(biofilm_excl, biofilm_imp)

# see limma user guide section 9.2 for more info about DE
# create design matrix
b_samp <- data.frame(T = (rep(c("EV", "WCL"), each = 5)))

b_design <- stats::model.matrix(~ 0 + T, data = b_samp)
colnames(b_design) <- c("EV", "WCL")

# define sample comparisons of interest
b_contrasts <- c("EV - WCL")

# make all pair-wise comparisons between EV and WCL
# and perform limma::eBayes()
b_efit <- limma_eBayes(biofilm_de, design = b_design, contrasts = b_contrasts)

# extract DE results
biofilm_res <- get_results(efit = b_efit, 
                           mat = biofilm_de, 
                           p_val = 0.01, 
                           lfc = 0, 
                           type = "individual")[[1]]
```

```{r atcc, include=FALSE}
# filter out reverse, contaminant proteins, proteins with <2 unique peptides
# extract LFQ intensity columns and UniProt accessions 
# log2 transform LFQ intensities
atcc_lfq <- atcc %>% 
  filter(Reverse != "+", 
         Potential.contaminant != "+", 
         Unique.peptides >= 2) %>% 
  convert_lfq(., atcc_exp)

# filter for proteins quantified in min 2/3 reps of: 
# A1 EV or A1 WCL or A9 EV or A9 WCL
atcc_filt <- filter_na4(atcc_lfq, "or", "<=", 
                        pat1 = "A10231_EV", 1, 
                        pat2 = "A10231_W", 1, 
                        pat3 = "A90028_EV", 1, 
                        pat4 = "A90028_W", 1)

# normalise LFQ intensities
atcc_norm <- normalizeCyclicLoess(atcc_filt)

# filter for proteins identified in min 1/3 reps of 10231 EV
# define 'ATCC10231 EV proteins'
atcc1_ev <- filter_na(atcc_norm, op = "<=", 
                      pat = "A10231_EV", val = 2)

# filter for proteins identified in min 1/3 reps of 10231 WCL
# define 'ATCC10231 WCL proteins'
atcc1_w <- filter_na(atcc_norm, op = "<=", 
                     pat = "A10231_W", val = 2)

# filter for proteins identified in min 1/3 reps of 90028 EV
# define 'ATCC90028 EV proteins'
atcc9_ev <- filter_na(atcc_norm, op = "<=", 
                      pat = "A90028_EV", val = 2)

# filter for proteins identified in min 1/3 reps of 90028 WCL
# define 'ATCC90028 WCL proteins'
atcc9_w <- filter_na(atcc_norm, op = "<=", 
                      pat = "A90028_W", val = 2)

# can't impute with so many missing values
# filter for exclusive to 10231 EV or WCL or 90028 EV or WCL (don't impute)
atcc_excl <- filter_na4(atcc_norm, logic = "or", op = "==", 
                        pat1 = "A10231_EV", val1 = 3, 
                        pat2 = "A10231_W", val2 = 3, 
                        pat3 = "A90028_EV", val3 = 3, 
                        pat4 = "A90028_W", val4 = 3)

# filter for proteins with min 1 valid value in all 4 sample types (impute)
atcc_both <- filter_na4(atcc_norm, logic = "and", op = "<=", 
                        pat1 = "A10231_EV", val1 = 2, 
                        pat2 = "A10231_W", val2 = 2, 
                        pat3 = "A90028_EV", val3 = 2, 
                        pat4 = "A90028_W", val4 = 2)

# proteins with missing values tend to have lower intensity
# therefore proteins are MNAR, close to detection limit
# use left censored imputation method
atcc_imp <- impute_QRILC(atcc_both)

# recombine imputed proteins and non-imputed proteins in matrix
atcc_de <- rbind(atcc_excl, atcc_imp)
colnames(atcc_de) <- stringi::stri_replace_all_regex(colnames(atcc_de), 
                                                     c("A10231_", "A90028_"), 
                                                     c("A1_", "A9_"), 
                                                     vectorize_all = FALSE)

# see limma user guide section 9.2 for more info about DE
# create design matrix
atcc_samp <- data.frame(T = (rep(c("A1_EV", "A1_WCL", 
                                   "A9_EV", "A9_WCL"), each = 3)))

atcc_design <- stats::model.matrix(~ 0 + T, data = atcc_samp)
colnames(atcc_design) <- c("A1_EV", "A1_W", "A9_EV", "A9_W")

# define sample comparisons of interest
atcc_contrasts <- c("A1_EV - A1_W", "A9_EV - A9_W")

# make all pair-wise comparisons for specified contrasts
# and perform limma::eBayes()
atcc_efit <- limma_eBayes(atcc_de, atcc_design, atcc_contrasts)

# extract overall results
atcc_overall <- get_results(efit = atcc_efit, mat = atcc_de, 
                            p_val = 0.01, lfc = 0, type = "overall")

# extract DE results
a1_res <- get_results(efit = atcc_efit, 
                      mat = atcc_de, 
                      p_val = 0.01, 
                      lfc = 0, 
                      type = "individual")[[1]]

a9_res <- get_results(efit = atcc_efit, 
                      mat = atcc_de, 
                      p_val = 0.01, 
                      lfc = 0, 
                      type = "individual")[[2]]
```

# Supplementary data

\beginsupplement

```{r figureS1, fig.cap=fS1_cap, fig.height=5, fig.width=5}
fS1_cap <- c("\\textbf{Size distribution of \\textit{C. albicans} EVs as 
              determined by nanoparticle tracking analysis (NTA).} A line plot 
              for each EV biological replicate from each strain is shown. Each 
              line plots represents the average of three technical replicates 
              and the dashed grey lines indicate 100 nm.")

figS1_labels <- c("ATCC10231", 
                  "ATCC90028", 
                  "DAY286 biofilm", 
                  "DAY286 yeast")

names(figS1_labels) <- c("ATCC10231", 
                         "ATCC90028", 
                         "Biofilm", 
                         "DAY286")

# average the nta technical replicate data
nta_techavg <- nta %>% 
  nanolyze(particle_size, Strain, Bio_rep, Dilution, 
           name = "Tech_rep", 
           param_var = True_count)

nta_techavg$Strain <- factor(nta_techavg$Strain, levels = c("DAY286",
                                                            "ATCC90028", 
                                                            "ATCC10231", 
                                                            "Biofilm"))

# plot nta distribution for each bio rep
nta_techavg %>% 
  ggplot(aes(x = particle_size, y = Tech_rep_mean, colour = Bio_rep)) + 
  geom_line(size = 0.5) + 
  geom_vline(xintercept = 100, colour = "#999999", linetype = "dashed") +
  facet_wrap(~ Strain, scales = "free", ncol = 2, nrow = 2, 
             labeller = labeller(Strain = figS1_labels)) + 
  theme_classic() + 
  theme(strip.background = element_blank(), 
        legend.position = "bottom", 
        plot.margin = margin(t = 0, l = 0, r = 0.1, b = 0, unit = "inches")) + 
  xlab(label = "Particle size (nm)") + 
  ylab(label = "Particles / mL") + 
  labs(colour = "Biological replicate")
```

\newpage

```{r tableS1Data}
# construct data.frame with nta experiment info
nta_metadata <- data.frame(id = c("ATCC10231_1", "ATCC10231_2", "ATCC10231_3", 
                                  "ATCC90028_1", "ATCC90028_2", "ATCC90028_3", 
                                  "Biofilm_1", "Biofilm_2", "Biofilm_3", 
                                  "Biofilm_4", "Biofilm_5", 
                                  "DAY286_1", "DAY286_2", "DAY286_3"), 
                           strain = c(rep("ATCC10231", 3), 
                                      rep("ATCC90028", 3), 
                                      rep("Biofilm", 5), 
                                      rep("DAY286", 3)), 
                           replicate = c(seq(1, 3), seq(1, 3), 
                                         seq(1, 5), seq(1, 3)), 
                           OD600 = c(0.455*20, 0.557*20, 0.615*20, 
                                     0.310*50, 0.337*50, 0.297*50, 
                                     rep(NA, 5), 
                                     36.2, 39.7, 34.0),
                           culture_vol = c(rep(150, 3),
                                           rep(150, 3), 
                                           rep(1000, 5), 
                                           rep(300, 3)), 
                           protein_yield = c(19.2, 28.2, 28.7, 
                                             42.4, 71.6, 58.4, 
                                             20.4, 19.8, 29.7, 19.3, 41.6, 
                                             21.9, 18.0, 22.4), 
                           protein_conc = c(0.192, 0.282, 0.287, 
                                            0.424, 0.716, 0.584, 
                                            0.204, 0.198, 0.297, 0.193, 0.416, 
                                            0.437, 0.360, 0.447), 
                           df = c(250, 750, 750, 
                                  1500, 3000, 3000, 
                                  3000, 2000, 2000, 3000, 6000, 
                                  1500, 1500, 1500), 
                           mean = c(248.5, 218.7, 217.9, 
                                    197.1, 233.1, 191.0, 
                                    119.4, 142.1, 143.3, 133.3, 120.9, 
                                    182.8, 174.5, 164.5), 
                           mode = c(171.9, 155.5, 140.2, 
                                    135.4, 156.7, 138.0, 
                                    100.5, 94.7, 103.1, 103.1, 97.5, 
                                    106.4, 99.2, 100.2), 
                           sd = c(120.1, 112.1, 110.0, 
                                  87.1, 90.7, 68.4, 
                                  47.4, 64.3, 70.2, 60.7, 47.4, 
                                  93.4, 93.9, 83.1), 
                           d10 = c(139.1, 110.2, 118.1, 
                                   122.4, 141.9, 126.0, 
                                   90.2, 88.8, 96.4, 93.2, 89.7, 
                                   97.8, 93.8, 93.0), 
                           d50 = c(211.9, 188.2, 180.3, 
                                   169.4, 209.3, 170.7, 
                                   105.6, 115.0, 110.9, 108.1, 105.9, 
                                   152.8, 147.3, 136.2),
                           d90 = c(393.0, 354, 354.4, 
                                   301.4, 381.1, 285.8, 
                                   151.5, 239.7, 240.8, 210.2, 171.2, 
                                   320.4, 308.0, 290.9),
                           sort_order = c(rep(3, 3), rep(2, 3), 
                                          rep(4, 5), rep(1, 3)), 
                           stringsAsFactors = FALSE)

# determine the total particle concentration for each biological replicate
nta_total <- nta_techavg %>% 
  nanocount(Strain, Bio_rep, Dilution, 
            param_var = Tech_rep_mean) %>% 
  tidyr::unite("id", Strain, Bio_rep) %>% 
  inner_join(., nta_metadata, by = "id") %>% 
  arrange(sort_order) %>% 
  select(strain, replicate, OD600, culture_vol, protein_yield, protein_conc, 
         Particle_count, mean, mode, sd) %>% 
  mutate(OD600 = case_when(is.na(OD600) == TRUE ~ "NA", 
                           TRUE ~ as.character(OD600)), 
         Particle_count = signif(Particle_count, digits = 3))
```

```{r tableS1}
tabS1_labs <- c("Biological\nreplicate", 
                "OD600", 
                "Culture vol.\n(mL)", 
                "Protein yield\n($\\mu$g)", 
                "Protein conc.\n($\\mu$g/mL)", 
                "Particle conc.\n(particles/mL)", 
                "Mean", 
                "Mode", 
                "SD")

tabS1_cap <- "\\textbf{Metadata for each EV isolation analysed in this study.} 
              The optical density (at 600 nm) and volume of each culture from 
              which the EVs were isolated is given. The protein and particle 
              concentration of the EVs along with the NTA size distribution 
              results for each biological replicate are also shown."

# construct a table of metadata for each EV biological replicate
nta_total %>% 
  select(-strain) %>% 
  mutate_all(linebreak) %>% 
  knitr::kable(caption = tabS1_cap, 
               format = "latex", 
               booktabs = TRUE, 
               linesep = "", 
               col.names = linebreak(tabS1_labs), 
               escape = FALSE, 
               align = "l") %>% 
  kable_styling(latex_options = c("scale_down", 
                                  "HOLD_position")) %>% 
  add_header_above(c(" " = 6, "Particle diameter (nm)" = 3)) %>% 
  pack_rows("DAY286 yeast", 1, 3) %>% 
  pack_rows("ATCC90028", 4, 6) %>% 
  pack_rows("ATCC10231", 7, 9) %>% 
  pack_rows("DAY286 biofilm", 10, 14)
```

\newpage

```{r tableS2Data}
# make list of WCL sig. enriched and exclusive proteins for each strain
wcl_comp <- list("yeast" = yeast_res %>% 
                   filter(group == "WCL up" | group == "WCL ex") %>% 
                   pull(CGD_gene_name), 
                 "biofilm" = biofilm_res %>% 
                   filter(group == "WCL up" | group == "WCL ex") %>% 
                   pull(CGD_gene_name), 
                 "ATCC10231" = a1_res %>% 
                   filter(group == "A1_W up" | group == "A1_W ex") %>% 
                   pull(CGD_gene_name), 
                 "ATCC90028" = a9_res %>% 
                   filter(group == "A9_W up" | group == "A9_W ex") %>% 
                   pull(CGD_gene_name))

# get venn segments
wcl_comp_tab <- plot_venn(vlist = wcl_comp, 
                          use_uniprot = FALSE, 
                          type = "df")

# extract protein names from venn overlap
wcl_candidates <- wcl_comp_tab %>% 
  filter(apply(.[, 1:4], 1, sum) == 4) %>% 
  pull(..values..) %>% 
  unlist() %>% 
  as.data.frame(stringsAsFactors = FALSE) %>% 
  rename(Name = 1)

# construct table S2
wcl_longlist <- wcl_candidates %>% 
  mutate(Function = match_id(Name, uniprot, "CGD_gene_name", 
                             "Protein_name", concat = TRUE, str_split = FALSE),
         DAY.Y = match_id(Name, yeast_res, "CGD_gene_name", 
                          "logFC", concat = TRUE, str_split = FALSE), 
         A9 = match_id(Name, a9_res, "CGD_gene_name", 
                       "logFC", concat = TRUE, str_split = FALSE), 
         A1 = match_id(Name, a1_res, "CGD_gene_name", 
                       "logFC", concat = TRUE, str_split = FALSE), 
         DAY.B = match_id(Name, biofilm_res, "CGD_gene_name", 
                          "logFC", concat = TRUE, str_split = FALSE), 
         TM = match_id(Name, uniprot, "CGD_gene_name", "Transmembrane"), 
         SP = match_id(Name, uniprot, "CGD_gene_name", "Signal_peptide")) %>% 
  mutate_at(.vars = vars(DAY.Y, DAY.B, A1, A9), 
            .funs = list(~ sprintf(as.numeric(.), fmt = "%#.2f"))) %>% 
  mutate_at(.vars = vars(DAY.Y, DAY.B, A1, A9), 
            .funs = list(~ case_when(. == "NA" ~ "ex", 
                                     TRUE ~ as.character(.)))) %>% 
  mutate(TM = stringr::str_count(TM, "Helical"), 
         SP = case_when(grepl("SIGNAL", SP) ~ "Y", 
                        TRUE ~ ""), 
         TM = case_when(TM == 0 ~ "", 
                        TRUE ~ as.character(TM))) %>% 
  rename("DAY Y" = DAY.Y, "DAY B" = DAY.B) %>% 
  arrange(Name)

# load vector of protein functions from external R script
<<fun_list2>>

# add proper protein functions to table 1
# also add vesicle depleted supernatant info and 
# group proteins according to localisation
wcl_longlist <- wcl_longlist %>% 
  mutate(Function = function_desc2, 
         VDM = c(rep("", 52)), 
         pack_row = c(8, rep(4, 2), 3, rep(2, 2), 7, rep(1, 2), rep(7, 3), 1, 
                      rep(2, 3), 7, 2, 4, rep(2, 2), 7, 3, 2, 5, rep(1, 2), 6, 
                      2, 7, 6, 2, 1, 7, rep(1, 3), 2, 1, 5, 2, 1, 6, 
                      1, 6, 2, 1, 9, 9, rep(1, 3))) %>% 
  arrange(pack_row)
```

(ref:table1) [@gil-bona2015]

(ref:footnote1) [@skrzypek2017]

(ref:footnote2) [@almagroarmenteros2017]

(ref:footnote3) [@savojardo2018]

```{r tableS2}
tabs2_cap <- "\\textbf{Candidate negative protein markers for 
              \\textit{C. albicans} EVs.} This list of proteins consists of 
              those that were found to be exclusive to whole cell lysate (WCL) 
              or significantly enriched in WCL across the four 
              \\textit{C. albicans} strains examined in this study. Proteins 
              are grouped according to their subcellular localisation as 
              annotated in the Candida Genome Database (candidagenome.org) 
              (ref:footnote1). The log\\textsubscript{2} ratio of the abundance 
              (mean MaxQuant LFQ intensity) of each protein in EVs compared to 
              WCL for each strain is listed. A negative value indicates that a 
              protein was enriched in WCL compared to EV. \`\`ex\" indicates 
              where a protein was only quantified in the WCL fraction and not 
              in EVs for that strain. The \`\`TM\" column indicates the number 
              of transmembrane domains for each protein as annotated in 
              UniProtKB. \`\`SP\" indicates whether a protein is annotated as 
              having a signal peptide according to UniProtKB. \`\`VDM\" shows 
              whether a protein has previously been detected in 
              vesicle-depleted culture media (i.e. the proteins may also be in 
              the soluble secretome) (ref:table1). Underlined proteins are 
              those identified as the best candidates for negative
              EV markers according to the criteria depicted in Figure 1."

# draw table S2
knitr::kable(wcl_longlist[, -10], 
             caption = tabs2_cap, 
             format = "latex", 
             booktabs = TRUE, 
             longtable = TRUE, 
             linesep = "", 
             escape = FALSE) %>% 
  row_spec(row = c(23, 30, 34, 35, 44, 50), underline = TRUE) %>% 
  add_header_above(c(" " = 2, 
                     "log\\\\textsubscript{2}(fold change) EV vs WCL" = 4, 
                     " " = 3), escape = FALSE) %>% 
  kable_styling(font_size = 8, 
                latex_options = c("HOLD_position")) %>% 
  pack_rows("Cytoplasm", 1, 16) %>% 
  pack_rows("Mitochondria", 17, 30) %>% 
  pack_rows("Nucleus", 31, 32) %>% 
  pack_rows("Vacuole", 33, 35) %>% 
  pack_rows("Peroxisome", 36, 37) %>% 
  pack_rows("Proteasome", 38, 41) %>% 
  pack_rows("Cell wall, cell surface, fungal biofilm matrix", 42, 49) %>% 
  pack_rows("Cytoskeleton", 50, 50) %>% 
  pack_rows("Ribosome", 51, 52) %>% 
  footnote(alphabet = c("Protein localisation was inferred from sequence 
                         similarity with \\\\textit{S. cerevisiae} ortholog as 
                         annotated in the Candida Genome Database 
                         (ref:footnote1)", 
                        "Protein localisation was predicted using DeepLoc-1.0 
                         (ref:footnote2)", 
                        "Protein localisation was predicted using BUSCA 
                         (ref:footnote3)"), 
           escape = FALSE, 
           threeparttable = TRUE)
```

\newpage

# References

<div id="refs"></div>