### 2019-06-18 README.Rmd
Adjust main README.md

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>")
```

# candidaev

![GitHub last commit](https://img.shields.io/github/last-commit/csdaw/candidaev.svg?style=popout)

This repository contains the research compendium for our paper:

> Dawson, C. S., Garcia-Ceron, D.,. Rajapaksha, H., Faou, P., Anderson, M. A., 
Bleackley, M. R. (2019) Sur7 family proteins are putative markers for
_Candida albicans_ extracellular vesicles. Mol. Cell. Proteomics, submitted.

It has been permanently archived at Zenodo with the DOI shown above.

The compendium includes all the data, code, and text associated with the 
publication. It is structured as an R package to facilitate reproducilibity and 
distribution. The principles and motivations behind using research compendiums 
for academic research is detailed in 
[Marwick et al. (2018)](https://doi.org/10.1080/00031305.2017.1375986).

The R package [rrtools](https://github.com/benmarwick/rrtools) was used in the 
production of this compendium.

### Citation

Please cite this compendium as:

> Dawson, C. S., Garcia-Ceron, D.,. Rajapaksha, H., Faou, P., Anderson, M. A., 
Bleackley, M. R. (2019) Research compendium for: _Sur7 family proteins are_
_putative markers for_ Candida albicans _extracellular vesicles_, 2019. 
Version x.x.x. Zenodo. URL.

### Contents

```
    .
    ├── /R                   # Script files for functions and included data
    ├── /data-raw            # Raw data used (but not modified) in analyses
    ├── /data                # Data tables installed with package
    ├── /inst/manuscript     # Final submitted manuscript and supplemental files
    ├── /man                 # Auto-generated documentation for functions and included data
    ├── /vignettes           # Detailed examples of how analyses were performed
    ├── .Rbuildignore        # List of files and directories to ignore during R package build
    ├── .gitattributes       # List of files for github/linguist to ignore
    ├── .gitignore           # List of files to ignore during git commit
    ├── DESCRIPTION          # Research compendium metadata
    ├── LICENSE.md           # MIT License for code
    ├── NAMESPACE            # Auto-generated file for function export
    ├── README.Rmd           # R markdown file to produce README.md
    ├── README.md            # Description of research compendium
    └── candidaev.Rproj      # R project file for compendium

```

### Accessing the compendium

#### 1. Downloading directly from GitHub
You can download the compendium as a zip from from this URL: 
<https://github.com/csdaw/candidaev/archive/master.zip>. 
Note that the contents of `/data-raw` is not installed when the package is 
installed but instructions for downloaded just these files is provided
[here](https://github.com/csdaw/candidaev/tree/master/data-raw).

#### 2. Cloning via `git`

#### 3. Installing the package using R and `devtools`

This research compendium can be installed as an R package `candidaev` from 
GitHub using `devtools`:
```{r gh-install, eval = FALSE}
# install.packages("devtools")
devtools::install_github("csdaw/candidaev")
```

### Using the `candidaev` package

### R session information

The system on which this document was compiled was running 
pandoc v`r rmarkdown::pandoc_version()`. Here is output of 
`sessionInfo()`.

```{r sessionInfo, echo=FALSE}
sessionInfo()
```