### 2019-07-03 R/README.Rmd
Update other README.md

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>")
```

# R

## Description

This directory contains R script files used throughout the project. Each file 
contains one custom function or one data.frame which has been
documented using an roxygen2 skeleton.

After installing `candidaev`, an index of all the included 
functions/data.frames can be viewed in R using:

```{r, eval=FALSE}
help(package = "candidaev")
```