### 2019-07-03 data-raw/README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>")
```

# /data-raw

## Description
This directory contains all of the raw data files used in this project. 
These files are *not* included when the package is installed using 
`devtools::install_github()`. Rather *'clean'* versions of the `.csv` and 
`.tab` files are provided in the [`/data`](../data) folder as `.rda` files 
which are accessible once the package has been installed and loaded or attached.

Some different ways to access these raw files include:

1. Right click on a desired file and select `Save link as...`.

2. Copy the url for the desired directory, paste into 
[DownGit](https://minhaskamal.github.io/DownGit/#/home), then 
download as a `.zip`.

3. Download this entire repository as a `.zip` from this URL: 
<https://github.com/csdaw/candidaev/archive/master.zip>

4. Clone this repository using `git`.

## Contents
```
    .
    ├── ...
    ├── /data-raw
    │    ├── /lcms           # proteinGroups.txt files output by MaxQuant
    │    ├── /nta            # Nanosight NS300 .csv summary tables
    │    ├── /reference      # UniProt and Candida Genome Database reference tables
    │    ├── README.Rmd      # R markdown file to produce README.md
    │    └── README.md       # Description of /data-raw directory
    └── ...

```