### 2019-07-03 inst/manuscript/README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>")
```

# /inst/manuscript

## Description

This directory contains the submitted manuscript, supplemental figures and 
tables, and supplemental files for our paper:

> Dawson, C. S., Garcia-Ceron, D., Rajapaksha, H., Faou, P., Anderson, M. A., 
Bleackley, M. R. (2019) Protein markers for _Candida albicans_ EVs include 
claudin-like Sur7 family proteins and Rho GTPases. Mol. Cell. Proteomics, 
submitted.

The Rmarkdown (`.Rmd`), `.svg`, and `.R` files used to generate these 
documents are provided along with the derived `.pdf` and `.csv` files.

## Contents
```
    .
    ├── ...
    ├── /inst
    │    ├── /manuscript
    │         ├── /components                 # Contains required components for building the manuscript
    │         ├── /figures                    # Contains .svg and .csv files used to make some figures
    │         ├── /supplement_files           # Contains .R scripts used to generate supplemental files
    │         ├── manuscript.Rmd              # R markdown file to produce manuscript.pdf
    │         ├── manuscript.pdf              # Final submitted manuscript
    │         ├── manuscript_supplement.Rmd   # R markdown file to produce manuscript_supplement.pdf
    │         ├── manuscript_supplement.pdf   # Final submitted supplemental figures and tables
    │         ├── README.Rmd                  # R markdown file to produce README.md
    │         └── README.md                   # Description of /inst/manuscript directory
    └── ...

```