### 2020-04-10 inst/manuscript/README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>")
```

# inst/manuscript

## Description

This directory contains the submitted manuscript and supplement `.docx` files, 
main figures, supplementary figures, supplementary data for our paper:

> Dawson, C. S., Garcia-Ceron, D., Rajapaksha, H., Faou, P., Anderson, M. A., 
Bleackley, M. R. (2020) Protein markers for _Candida albicans_ EVs include 
claudin-like Sur7 family proteins. _Journal of Extracellular_ 
_Vesicles_, x(x): 1750810. DOI: 
[10.1080/20013078.2020.1750810](https://doi.org/10.1080/20013078.2020.1750810)

The scripts used to generate the figures, tables, and supplementary data are
also present. This directory is automatically installed when the `candidaev`
package is installed.

## Contents
```
.
├── ...
├── inst
│    ├── manuscript
│         ├── components                        # Required components for building the manuscript
│         ├── figures                           # Individual high resolution files for each figure
|         ├── tables                            # Tables as .docx files
│         ├── supplementary_data                # Supplementary data and scripts to generate them
│         ├── README.Rmd                        # Rmarkdown file to produce README.md
│         ├── README.md                         # Description of inst/manuscript directory
│         ├── candidaev_manuscript_FINAL.docx   # Manuscript as accepted
│         ├── candidaev_supplement_FINAL.docx   # Supplement figure and table legends
│         ├── candidaev_figures.pdf             # Main figures and table as accepted
│         ├── candidaev_figures.Rmd             # Script to produce main figures and table
│         ├── candidaev_supplement.pdf          # Supplementary figures and tables as accepted
│         └── candidaev_supplement.Rmd          # Script to produce supplementary figures and tables
└── ...

```